---
title: "CollegeBasketballData (CBD) Endpoints"
description: "A tour of the <code>cbbd_*</code> functions, which wrap the CollegeBasketballData API (api.collegebasketballdata.com) for men's college basketball games, plays, stats, ratings, recruiting and more."
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/hoopR-data/blob/main/themes/hoopR_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The `cbbd_*()` functions wrap the [CollegeBasketballData API](https://collegebasketballdata.com),
the men's college basketball sibling of CollegeFootballData (which `cfbfastR`
wraps with `cfbd_*`). They cover the full v1 endpoint surface — games, play-by-play,
substitutions, team/player statistics, lineups, ratings, rankings, betting lines,
recruiting, the transfer portal, the NBA draft, and reference data for teams,
venues and conferences.

Every wrapper returns a tidy `hoopR_data` tibble and prints its usage with the
shared hoopR theming.

## Registering an API key

The CollegeBasketballData API requires a free Bearer-token API key. Request one at
[collegebasketballdata.com/key](https://collegebasketballdata.com/key), then store
it as the `CBBD_API_KEY` environment variable. The easiest way is to add it to your
`.Renviron` file:

```{r}
usethis::edit_r_environ()
# add this line, save, and restart R:
# CBBD_API_KEY = XXXX-YOUR-API-KEY-HERE-XXXXX
```

For one-off use within a single session:

```{r}
Sys.setenv(CBBD_API_KEY = "XXXX-YOUR-API-KEY-HERE-XXXXX")
```

Confirm the key is detected and inspect the registration help:

```{r}
library(hoopR)

has_cbbd_key()
#> [1] TRUE

?register_cbbd
```

## Reference data: conferences, teams, venues

Start with the lookups you will use to filter the other endpoints. The `id` values
returned here are the CollegeBasketballData ids consumed elsewhere; `source_id`
holds the matching ESPN id.

```{r}
# Conferences (and historical membership)
cbbd_conferences()
cbbd_conferences_history(conference = "ACC")

# Teams for a season, and a single team's roster
cbbd_teams(conference = "ACC", season = 2024)
cbbd_teams_roster(season = 2024, team = "Duke")

# Venues
cbbd_venues()
```

## Games and box scores

`cbbd_games()` returns the schedule/results; the `*_teams` and `*_players` variants
return team- and player-level box scores. `cbbd_scoreboard()` returns live and
same-day games (so it is empty in the off-season).

```{r}
games <- cbbd_games(season = 2024, team = "Duke")
games

cbbd_games_media(season = 2024, team = "Duke")    # broadcast info
cbbd_games_teams(season = 2024, team = "Duke")    # team box scores
cbbd_games_players(season = 2024, team = "Duke")  # player box scores

cbbd_scoreboard(conference = "ACC")
```

Most by-id endpoints take a CollegeBasketballData `game_id` from `cbbd_games()`:

```{r}
game_id <- games$id[1]
```

## Play-by-play

Pull play-by-play by game, by player, by team, by date, or by tournament. Pass
`shooting_plays_only = TRUE` to restrict to shot events.

```{r}
cbbd_plays_game(game_id = game_id)
cbbd_plays_team(season = 2024, team = "Duke")
cbbd_plays_player(player_id = 160, season = 2024)
cbbd_plays_date(date = "2024-02-01T00:00:00.000Z")
cbbd_plays_tournament(tournament = "NCAA", season = 2024)

# Reference: the list of play types
cbbd_play_types()
```

## Substitutions

```{r}
cbbd_substitutions_game(game_id = game_id)
cbbd_substitutions_team(season = 2024, team = "Duke")
cbbd_substitutions_player(player_id = 160, season = 2024)
```

## Team and player statistics

Season aggregates for teams and players, plus dedicated shooting splits and a
team leaderboard. Nested statistic objects are flattened into prefixed columns
(e.g. `team_stats_*`, `field_goals_*`).

```{r}
cbbd_stats_team_season(season = 2024, team = "Duke")
cbbd_stats_team_leaderboard(season = 2024)
cbbd_stats_team_shooting_season(season = 2024, team = "Duke")

cbbd_stats_player_season(season = 2024, team = "Duke")
cbbd_stats_player_shooting_season(season = 2024, team = "Duke")
```

## Lineups

Five-player lineup stats for a team-season or a single game.

```{r}
cbbd_lineups_team(season = 2024, team = "Duke")
cbbd_lineups_game(game_id = game_id)
```

## Ratings and rankings

```{r}
cbbd_ratings_srs(season = 2024)        # Simple Rating System
cbbd_ratings_adjusted(season = 2024)   # adjusted offensive/defensive efficiency
cbbd_ratings_elo(season = 2024)        # Elo

cbbd_rankings(season = 2024, poll_type = "ap")
```

## Betting lines

```{r}
cbbd_lines(season = 2024, team = "Duke")
cbbd_lines_providers()
```

## Recruiting and the transfer portal

```{r}
cbbd_recruiting_players(year = 2024)
cbbd_recruiting_teams(year = 2024)
cbbd_recruiting_portal(year = 2024)
```

## NBA draft

```{r}
cbbd_draft_teams()
cbbd_draft_positions()
cbbd_draft_picks(year = 2024)
```

## Where to go next

- Browse the full function reference under the **CollegeBasketballData (CBD)**
  section of the [reference index](../reference/index.html).
- Each wrapper's help page (`?cbbd_games`, `?cbbd_plays_game`, ...) documents its
  parameters and the columns it returns.
- Inspect your remaining API usage and rate limits in the response headers from
  the [CollegeBasketballData API](https://collegebasketballdata.com).
