---
title: "Intro to kenpomR"
author: "Saiem Gilani"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to kenpomR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
old <- options(rmarkdown.html_vignette.check_title = FALSE)
```

We will be acquiring data from [kenpom.com](https://kenpom.com), using the ```kenpomR``` package, created by [Saiem Gilani](https://twitter.com/saiemgilani). An active subscription to the website will be required for most of this tutorial.

## R & RStudio
This tutorial will require the use of R and RStudio. You can [follow the instructions at R Studio on how to get started](https://rstudio.com/products/rstudio/download/). 

## First, save your login information privately
You can use .Renviron to save your `kp_user` and `kp_pw` variables like so: 
```{r eval = FALSE}
file.edit("~/.Renviron")
```
Then save that information in a format like this:

```
kp_user = "your email here"
kp_pw = "your password here"

```
Then we can reference these in our scripts with `Sys.env("kp_user")` or `Sys.env("kp_pw")`.

## Import libraries and set browser login
```{r, warning=FALSE,  message = FALSE}
library(dplyr)
library(kenpomR)
browser <- login(Sys.getenv("kp_user"), Sys.getenv("kp_pw"))

```

```{r message = FALSE, results='hide'}
rtgs <- get_pomeroy_ratings(browser, min_year = 2006, max_year = 2020)

```
```{r}
glimpse(rtgs)
```



