---
title: "KenPom Example Functions"
description: "Some examples to get started with using <code>hoopR</code> and KenPom's men's college basketball website."
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/hoopR-data/blob/main/themes/hoopR_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

## Ratings Table Function examples
```{r eval = FALSE}
#---- kp_ratings_tables.R ----------------
pomeroy_ratings <- kp_pomeroy_ratings(min_year = 2020, max_year = 2020)

efficiency <- kp_efficiency(min_year = 2020, max_year = 2020)

fourfactors <- kp_fourfactors(min_year = 2020, max_year = 2020)

pointdist <- kp_pointdist(min_year = 2020, max_year = 2020)

height <- kp_height(min_year = 2020, max_year = 2020)

teamstats <- kp_teamstats(min_year = 2020, max_year = 2020)

playerstats <- kp_playerstats(metric = 'eFG', conf_only = FALSE, year = 2020)

kpoy <- kp_kpoy(year = 2020)

```


## Team Tables
```{r eval = FALSE}


#---- kp_team_tables.R -------------------
team_sched <- kp_team_schedule(team = 'Florida St.', year = 2020)

gameplan <- kp_gameplan(team = 'Florida St.', year = 2020)

opptracker <- kp_opptracker(team = 'Florida St.', year = 2020)

team_players <- kp_team_players(team = 'Florida St.', year = 2020)

minutes_matrix <- kp_minutes_matrix(team = 'Florida St.', year = 2020)

team_player_stats <- kp_team_player_stats(team = 'Florida St.', year = 2020)

team_player_stats1 <- team_player_stats[[1]]

team_player_stats2 <- team_player_stats[[2]]

team_depth_chart <- kp_team_depth_chart(team = 'Florida St.', year = 2020)

team_lineups <- kp_team_lineups(team = 'Florida St.', year = 2020)

```


## Historical Tables
```{r eval = FALSE}

#---- kp_historical_tables.R ------------
team_history <- kp_team_history(team = 'Florida St.')

coach_history <- kp_coach_history(coach = 'Leonard Hamilton')

program_ratings <- kp_program_ratings()

# records_team appears to be working but there isn't anything
# there anymore at the moment for all years
# records_team <- get_records_team(browser, team = 'Florida St.')

pomeroy_archive_ratings <- kp_pomeroy_archive_ratings(date = '2018-11-22')

conf <- kp_conf(year = '2020',conf = 'ACC')

conf_stats <- kp_confstats(year = '2020')

conf_history <- kp_confhistory(conf = 'ACC')

```


## Miscellaneous Stats Tables
```{r eval = FALSE}
#---- kp_misc_stats.R ------------------
trends <- kp_trends()


officials <- kp_officials(year = 2020)

hca <- kp_hca()


arenas <- kp_arenas(year = 2020)



game_attrs <- kp_game_attrs(year = 2020, attr = "Excitement")

fanmatch <- kp_fanmatch(date = "2020-03-10")


```

# **Our Authors**
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>

## **Our Contributors**
-   [Jason Lee](https://x.com/theFirmAISports)
<a href='https://x.com/theFirmAISports' target='blank'><img src='https://img.shields.io/twitter/follow/theFirmAISports?color=blue&amp;label=%40theFirmAISports&amp;logo=x&amp;style=for-the-badge' alt='@theFirmAISports'/></a>
-   [Billy Fryer](https://x.com/BillyFryer42)
<a href='https://x.com/BillyFryer42' target='blank'><img src='https://img.shields.io/twitter/follow/BillyFryer42?color=blue&amp;label=%40BillyFryer42&amp;logo=x&amp;style=for-the-badge' alt='@BillyFryer42'/></a>
-   [Ross Drucker](https://x.com/rossdrucker9)
<a href='https://x.com/rossdrucker9' target='blank'><img src='https://img.shields.io/twitter/follow/rossdrucker9?color=blue&amp;label=%40rossdrucker9&amp;logo=x&amp;style=for-the-badge' alt='@rossdrucker9'/></a>
-   [Vladislav Shufinskiy](https://x.com/vshufinskiy)
<a href='https://x.com/vshufinskiy' target='blank'><img src='https://img.shields.io/twitter/follow/vshufinskiy?color=blue&amp;label=%40vshufinskiy&amp;logo=x&amp;style=for-the-badge' alt='@vshufinskiy'/></a>

## **Citation**

To cite the [**`hoopR`**](https://hoopR.sportsdataverse.org/) R package in publications, use:

BibTex Citation

```bibtex
@misc{hoopr,
  author = {Saiem Gilani},
  title = {hoopR: The SportsDataverse's R Package for Men's Basketball Data},
  url = {https://hoopR.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
