---
title: "Men's College Basketball Cookbook"
description: "A storytelling tour of men's college basketball analysis with <code>hoopR</code> -- recipes from a single game to a full bracket, plus KenPom efficiency, and a running commentary on how to read the package's function names."
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/hoopR-data/blob/main/themes/hoopR_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = "#>", collapse = TRUE,
                      warning = FALSE, message = FALSE, error = TRUE)
```

## Before we cook

Men's college basketball is a sprawl -- 360-odd Division I teams, a
month of March that decides everything, and a stats culture built as
much on tempo-free efficiency as on raw box scores. `hoopR` covers all
of it, and the trick to not drowning is the same as in the NBA
cookbook: **learn the grammar of the function names and you can guess
your way to almost anything.**

If you've read the NBA cookbook, you already know most of the grammar.
This one adds one new prefix -- `kp_` for KenPom -- and swaps the league
token from `nba` to `mbb`. That's nearly the whole diff.

### The grammar, refreshed for college

A `hoopR` college function answers three questions in order:

1. **Where's the data from?** -- the prefix.
   - `espn_` -- ESPN's APIs. The broad, stable backbone.
   - `kp_` -- KenPom (kenpom.com). Tempo-free efficiency, the lingua
     franca of college analytics. **Requires a KenPom subscription.**
   - `load_` -- pre-built bulk season files.
2. **Which league?** -- for ESPN it's spelled out: `espn_mbb_`. The
   `mbb` token is "men's basketball." (KenPom is men's-college-only, so
   `kp_` functions don't need a league token -- the prefix already
   implies it.)
3. **What do you want?** -- the rest of the name, general to specific.

So "a team's schedule from ESPN" is `espn_` + `mbb_` + `team_schedule`
-> `espn_mbb_team_schedule()`. "KenPom's efficiency ratings" is `kp_` +
`efficiency` -> `kp_efficiency()`. You will be right far more often than
you'll be wrong, and being wrong just means opening the reference index.

```{r}
library(hoopR)
library(dplyr)
```

## Recipe 1: A blue-blood program's season

**The story.** It's November and you want to set the table for Duke's
season -- schedule, roster, and where they sit.

Type `espn_mbb_team` and the grammar lays out the menu for you:

```{r}
team_id <- 150            # Duke
season  <- most_recent_mbb_season()

espn_mbb_team(team_id = team_id)                              # identity + record
espn_mbb_team_schedule(team_id = team_id, season = season)    # the slate
espn_mbb_team_roster(team_id = team_id, season = season)      # the players
espn_mbb_team_season_statistics(team_id = team_id, season = season)
```

The `most_recent_mbb_season()` helper is worth a margin note. Every
league in `hoopR` has a `most_recent_<league>_season()` companion, and
the new `espn_mbb_*` functions use it as their default `season`
argument. So most of the time you can simply *omit* `season` and get
the current year -- the function already knows what "now" means.

## Recipe 2: The freshman phenom

**The story.** Every college season has a one-and-done freshman everyone
argues about. You want that player's numbers.

College players turn over fast -- this year's phenom is next year's
pro -- so the reliable move is to pull the current roster first and
lift an `athlete_id` straight off it, then walk the `athlete_` family:

```{r}
roster     <- espn_mbb_team_roster(team_id = 150, season = most_recent_mbb_season())
athlete_id <- roster$athlete_id[1]   # whoever is first on the roster

espn_mbb_player_info(athlete_id = athlete_id)            # bio
espn_mbb_player_gamelog(athlete_id = athlete_id, season = most_recent_mbb_season())
espn_mbb_player_career_stats(athlete_id = athlete_id)    # career rollup, long format
```

That "pull the index, lift an ID, drill down" move is the single most
useful habit in the whole package. `athlete_gamelog` in particular only
has data for players *currently* on a roster -- last year's star has
moved on -- so feeding it an ID you just lifted off `team_roster` is
how you guarantee it returns something.

Margin note: `athlete_info`, `athlete_gamelog`, `athlete_career_stats`
all share the `athlete_` stem. When a stem repeats across many
functions, that stem is a *namespace* -- a promise that everything
under it is about the same subject. Browsing by stem (`espn_mbb_athlete`
+ Tab) is often faster than searching.

## Recipe 3: Tempo-free, the KenPom way

**The story.** Box scores lie a little in college -- a team that plays
fast will pile up points without being good. You want *efficiency*:
points per possession, not per game. That's KenPom's whole reason to
exist.

This is where the new prefix earns its keep. Everything KenPom is `kp_`.

```{r eval = FALSE}
# KenPom needs a subscription. Log in once per session.
kp_user_email <- "you@example.com"
kp_password   <- Sys.getenv("KP_PASSWORD")

# The famous Pomeroy ratings -- adjusted efficiency for every team.
kp_pomeroy_ratings(min_year = 2024, max_year = 2025)

# The four factors: shooting, turnovers, rebounding, free throws.
kp_fourfactors(min_year = 2025, max_year = 2025)

# Raw adjusted-efficiency table.
kp_efficiency(min_year = 2025, max_year = 2025)
```

The grammar lesson here is about *prefixes as capabilities*. The moment
you see `kp_`, three things are implied without another word: the data
is men's-college-only, it is tempo-free, and it needs authentication.
A prefix is a compressed paragraph. Learn what each prefix promises and
half the documentation becomes redundant.

KenPom's own family is deep -- and, of course, regular:

```{r eval = FALSE}
kp_team_schedule(team = "Duke", year = 2025)
kp_teamstats(min_year = 2025, max_year = 2025)
kp_playerstats(metric = "eFG", year = 2025)
kp_kpoy(year = 2025)         # KenPom Player of the Year race
kp_fanmatch(date = "2025-02-01")   # every game on a date, win probs attached
```

## Recipe 4: Game flow without the bracket pressure

**The story.** A January road game went to overtime. You want the play-
by-play and a sense of the swing.

Same three zoom levels as the NBA cookbook -- the grammar is
league-agnostic:

```{r}
game_id <- 401256760

espn_mbb_game_all(game_id = game_id)      # everything, as a named list
espn_mbb_team_box(game_id = game_id)      # team box
espn_mbb_player_box(game_id = game_id)    # player box
espn_mbb_pbp(game_id = game_id)           # play-by-play

# ESPN's win-probability-per-play for the same game.
espn_mbb_game_probabilities(event_id = game_id)
```

If you read the NBA cookbook, you've now seen `espn_*_game_all`,
`espn_*_team_box`, `espn_*_player_box`, `espn_*_pbp` in two leagues with
*only the league token changing*. That's the payoff of a regular
grammar: a recipe you learned for the NBA transfers to college by
editing one word.

## Recipe 5: Per-player, per-game box scores

**The story.** You want a single player's line from a single game, tidy
and long -- not the whole `player_box`.

```{r}
# event + team + athlete -- three IDs, because you're naming one cell
# of one game. (More identifiers = finer grain. Always.)
espn_mbb_game_player_box(
  event_id   = 401256760,
  team_id    = 52,
  athlete_id = 4277850
)
```

## Recipe 6: The conference race

**The story.** It's February and your league is a three-team knife
fight. You want the standings and the structure underneath them.

College conferences are a *hierarchy* -- divisions, conferences, the
whole D-I grouping -- and ESPN models that hierarchy with the
`season_group` family.

```{r}
season <- most_recent_mbb_season()

# A grammar gotcha worth flagging: espn_mbb_standings() takes `year`,
# not `season`. Most functions say `season`, a few older ones say
# `year` -- when an argument name surprises you, ?function is the
# two-second fix.
espn_mbb_standings(year = season)

# The group tree: groups -> one group -> its teams.
espn_mbb_season_groups(season = season, season_type = 2)
espn_mbb_season_group(group_id = 50, season = season, season_type = 2)
espn_mbb_season_group_teams(group_id = 50, season = season, season_type = 2)
```

Margin note: `season_groups` (plural) is the index; `season_group`
(singular) is one item. That singular/plural pair shows up all over the
package -- `franchises`/`franchise`, `tournaments`/`tournament`,
`positions`/`position`. Plural = "give me the list," singular = "give me
this one." It's the most reliable two-word rule in the whole grammar.

## Recipe 7: March

**The story.** It's the only story in college basketball. You want the
tournament.

```{r}
season <- most_recent_mbb_season()

espn_mbb_tournaments()                                  # which tournaments exist
espn_mbb_tournament(tournament_id = 3)                  # one tournament
espn_mbb_tournament_seasons(tournament_id = 3)          # its yearly editions
espn_mbb_tournament_season(tournament_id = 3, season = season)
```

Same singular/plural rule as Recipe 6, same general-to-specific march
from `tournaments` down to one `tournament_season`. By now you should be
able to *predict* this family before you read it -- which is the entire
point of the cookbook.

## Recipe 8: A whole season, in bulk

**The story.** You're done with single games. You want every college
game of a season to model on.

```{r eval = FALSE}
# load / mbb / pbp -- one call, one season.
pbp <- load_mbb_pbp(seasons = 2024)

load_mbb_player_box(seasons = 2024)
load_mbb_team_box(seasons = 2024)
load_mbb_schedule(seasons = 2024)
```

And straight into a database, with the same `(seasons, dbConnection,
tablename)` plumbing every `load_` function shares:

```{r eval = FALSE}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "mbb.sqlite")
load_mbb_pbp(seasons = 2020:2024, dbConnection = con, tablename = "mbb_pbp")
dbDisconnect(con)
```

## Working through a proxy

Campus networks love a proxy. `hoopR` handles them in three layers --
reach for the least invasive one that works.

**Layer 1 -- set it once for the session.** Covers everything: ESPN,
KenPom, loaders.

```{r eval = FALSE}
options(hoopR.proxy = "http://proxy.university.edu:8080")

# Authenticated proxy? Pass a list instead of a string.
options(hoopR.proxy = list(
  url      = "http://proxy.university.edu",
  port     = 8080,
  username = "netid",
  password = Sys.getenv("PROXY_PASS")
))

# Then just work -- every call inherits it.
espn_mbb_team_roster(team_id = 150, season = most_recent_mbb_season())
kp_pomeroy_ratings(min_year = 2025, max_year = 2025)
```

**Layer 2 -- environment variables.** For shared scripts and CI, where
the proxy belongs to the machine and not the analysis, set the standard
variables and let libcurl pick them up:

```{r eval = FALSE}
Sys.setenv(
  http_proxy  = "http://proxy.university.edu:8080",
  https_proxy = "http://proxy.university.edu:8080",
  no_proxy    = "localhost,127.0.0.1"
)
```

A grammar-flavored note on what *doesn't* take a per-call `proxy =`
argument: the `espn_*`, `kp_*`, and `load_*` functions all call the HTTP
layer directly, so they read the proxy from the session option or the
environment -- not from an argument. (In the NBA cookbook, the `nba_*`
Stats API functions *do* accept per-call `proxy =`, because they thread
`...` through. College basketball has no equivalent NBA-Stats-style
surface, so for `mbb` you'll use Layer 1 or Layer 2 every time.) The
prefix, once again, tells you the capability before you go looking.

## Where to go next

The college game has more teams, a deeper conference hierarchy, and the
`kp_` prefix -- but the grammar is identical to the NBA's. You decided
where the data lived, you named the league, you named the thing from
general to specific, and you let the singular/plural rule and the
shared stems carry you the rest of the way.

The women's game runs on `wehoop`, a sibling package with the exact
same grammar. `espn_mbb_team_roster` has a mirror image in
`espn_wbb_team_roster`; everything you just learned crosses straight
over. That's the WBB cookbook.
