---
title: "NBA Cookbook"
description: "A storytelling tour of NBA analysis with <code>hoopR</code> -- recipes that move from a single box score to a full season, with an eye on how the package's function names are built so you can guess the next one before you look it up."
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/hoopR-data/blob/main/themes/hoopR_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>", collapse = TRUE)
```

## Before we cook

Every recipe in this cookbook is a small story: a question you might
actually ask about the NBA, and the shortest honest path from that
question to a data frame you can work with. None of the code blocks run
when this page is built -- they all reach out to live APIs -- but every
one of them works when you paste it into an R session.

Here's the one idea worth internalizing before anything else: **`hoopR`
function names are not arbitrary. They're a grammar.** Once you can read
the grammar, you can *guess* the function you need instead of hunting
for it. That guessing skill is the real subject of this cookbook; the
basketball is just the excuse.

### The grammar of a `hoopR` function name

Almost every function answers three questions in order:

1. **Where does the data come from?** -- the prefix.
   - `espn_` -- ESPN's APIs. Broadest coverage, most stable.
   - `nba_` -- the official NBA Stats API (stats.nba.com). Deepest box
     scores, tracking data, lineup data.
   - `kp_` -- KenPom (men's college only -- you won't use it here, but
     you'll meet it in the MBB cookbook).
   - `load_` -- pre-built bulk data. A whole season in one call.
   - `nbagl_` -- the NBA G League.
2. **Which league / which slice?** -- for ESPN functions the league is
   baked in: `espn_nba_`. For the NBA Stats API the `nba_` prefix *is*
   the league.
3. **What thing do you want?** -- the rest of the name, read left to
   right from general to specific: `team_season_roster`,
   `athlete_career_stats`, `event_competitor_linescores`.

So if you want "a team's roster for a given season, from ESPN," you can
assemble the name in your head before you ever open the docs:
`espn_` + `nba_` + `team` + `_roster` -> `espn_nba_team_roster()`. If
you want "the league leaders from NBA.com," it's `nba_` + `leagueleaders`
-> `nba_leagueleaders()`. The package has hundreds of functions; you do
not need to memorize them. You need to memorize the grammar.

We'll point at the grammar in the margins of each recipe, like a cooking
show host muttering "and notice we salt the water *first*."

```{r}
library(hoopR)
library(dplyr)
```

## Recipe 1: One game, three levels of detail

**The story.** A friend texts you about last night's game and you want
to actually look at it -- not a headline, the game itself.

ESPN gives you a game at three zoom levels, and the function names tell
you which zoom you're asking for. Start wide, then push in.

```{r}
game_id <- 401283399

# Widest: the whole game summary as a named list of tibbles.
# Read the name: espn / nba / "game" / "all" -- give me all of it.
full_game <- espn_nba_game_all(game_id = game_id)
names(full_game)

# Medium: the box score. "team_box" and "player_box" are siblings --
# same noun (box), different grain (team vs. player).
team_box   <- espn_nba_team_box(game_id = game_id)
player_box <- espn_nba_player_box(game_id = game_id)

# Tightest: the play-by-play. Every possession, one row each.
pbp <- espn_nba_pbp(game_id = game_id)
```

The lesson hiding in the margin: `team_box` and `player_box` differ by
exactly one word, and that word is the *grain* of the table. Whenever
you see two functions whose names differ by a single noun, that noun is
almost always telling you "one row per ___."

## Recipe 2: A superstar's whole career on one screen

**The story.** You want to see LeBron James's career arc -- not this
season, the whole thing -- and you want it tidy enough to plot.

The word you're reaching for is **career**. ESPN's core-v2 API has a
dedicated career rollup, and the function is named exactly what it is.

```{r}
# espn / nba / athlete / career / stats. The name is the sentence.
lebron <- espn_nba_athlete_career_stats(athlete_id = 1966)

# It comes back long -- one row per (stat_type x category x stat) --
# which is exactly the shape you want for a faceted ggplot.
lebron |>
  filter(category_name == "offensive", stat_name == "avgPoints") |>
  select(stat_type_id, value)
```

Margin note: `athlete_career_stats()` defaults to fetching *both*
regular season and postseason and stacking them, keyed by a
`stat_type_id` column. A lot of the newer `espn_nba_*` functions do
this -- when a function could sensibly return "regular + playoffs," it
usually returns both and lets you filter. If you only want one, pass
`season_type`.

Want the season-by-season index instead of the career rollup? The
grammar predicts it: swap `career_stats` for `seasons`.

```{r}
espn_nba_athlete_seasons(athlete_id = 1966)
```

## Recipe 3: Game flow and the comeback

**The story.** The game in Recipe 1 had a wild fourth quarter. You want
to *see* the momentum swing.

```{r}
# Route A -- ESPN's per-play win probability, ready to plot as-is.
wp <- espn_nba_event_probabilities(event_id = 401283399)

# Route B -- if you'd rather compute momentum yourself, the NBA Stats
# API's V3 play-by-play is the rawest material. The prefix flips to
# `nba_` -- different data source, different first token -- and it
# keys off stats.nba.com game ids, which look nothing like ESPN's, so
# lift a real one off the schedule first.
sched  <- nba_schedule()
pbp_v3 <- nba_playbyplayv3(game_id = sched$game_id[1])
```

This is the grammar doing real work. `espn_nba_pbp()` and
`nba_playbyplayv3()` answer almost the same question -- give me the
plays -- and you can tell at a glance that the first is ESPN's take and
the second is NBA.com's, *because the prefix is the data source*. When
two functions look like near-synonyms, check the prefix; that's the
tiebreaker. (A small caveat the grammar can't tell you: the NBA Stats
API churns faster than ESPN's. If a `nba_*` function ever tells you it
is deprecated, it will name its replacement in the same breath -- here,
the older `nba_winprobabilitypbp()` now points you at
`nba_playbyplayv3()`. Read the message; it is doing you a favor.)

## Recipe 4: How did each player do in one game, one at a time?

**The story.** You want one player's stat line from one game, broken
out into a tidy long table -- not the whole `player_box`, just them.

This is one of the newer core-v2 endpoints, and its name is a little
sentence: a player's *box* score for one *event*.

```{r}
# espn / nba / event / player_box. "event" = one game; "player_box" =
# one athlete's line. You need the (event, team, athlete) trio because
# that's exactly what uniquely identifies "this player in this game."
kyle_anderson <- espn_nba_event_player_box(
  event_id   = 401283399,
  team_id    = 29,
  athlete_id = 2993874
)
```

Margin note on *why three IDs*: a function's arguments are a hint about
its grain. One ID (`game_id`) gets you a whole game. Three IDs
(`event` + `team` + `athlete`) get you one cell of it. If a function
asks for more identifiers, it's because it returns something more
specific -- the arguments and the grain move together.

While you're in event-detail land, the siblings all follow the pattern
`espn_nba_event_*`:

```{r}
espn_nba_event_competitor_linescores(event_id = 401283399, team_id = 29)
espn_nba_event_officials(event_id = 401283399)
espn_nba_event_play(event_id = 401283399, play_id = 4012833994)
```

## Recipe 5: A team's season at a glance

**The story.** You're previewing a team and want the essentials --
record, schedule, roster, and how they actually stack up statistically.

Everything you need starts with `espn_nba_team`. Type that much and let
autocomplete show you the menu; the grammar guarantees the menu is
sensibly named.

```{r}
team_id <- 13   # Los Angeles Lakers
season  <- most_recent_nba_season()

# The "season profile" is a hub -- it hands you $refs to everything
# else about this team-season.
profile <- espn_nba_team_season_profile(team_id = team_id, season = season)

espn_nba_team_schedule(team_id = team_id, season = season)
espn_nba_team_roster(team_id = team_id, season = season)

# The full team-season stat sheet, with the team's league RANK baked
# into every row -- "team_season_statistics" is the long-format sheet,
# "team_record" is just the W-L.
espn_nba_team_season_statistics(team_id = team_id, season = season)
espn_nba_team_record(team_id = team_id, season = season)
```

Notice the family resemblance: `team_season_profile`,
`team_season_roster`, `team_season_statistics`. The shared middle
(`team_season`) is the subject; the last word is the lens. Find one
member of a family and you've found the whole family.

## Recipe 6: The MVP leaderboard race

**The story.** It's March and you want the scoring, rebounding, and
assist leaders -- the raw material of every MVP argument.

```{r}
# NBA.com's leaders. Prefix `nba_` -> this is the official Stats API.
nba_leagueleaders(season = "2024-25", stat_category = "PTS")

# ESPN's version of the same idea. Same question, different kitchen --
# and the prefix tells you which kitchen before you read another word.
espn_nba_leaders(season = most_recent_nba_season())
```

If you ever forget whether leaders live under `nba_` or `espn_nba_`:
they live under *both*, because both APIs publish them. That's normal.
The package mirrors the data sources honestly rather than papering one
over the other -- so you choose based on which columns you want, not on
which name you happened to remember.

## Recipe 7: Tracking a draft class

**The story.** You want to follow a draft class -- who went where, and
in what order.

The draft has a little family of its own, and the names sort from
general to specific exactly as the grammar promises:

```{r}
season <- 2024

espn_nba_season_draft(season = season)        # the draft, top level
espn_nba_draft_rounds(season = season)        # broken out by round
espn_nba_draft_athletes(season = season)      # every drafted player
espn_nba_draft_athlete_detail(season = season, athlete_id = 108206)
espn_nba_draft_pick(season = season, round = 1, pick = 1)
```

Read that list top to bottom: `season_draft` -> `draft_rounds` ->
`draft_athletes` -> `draft_athlete_detail` -> `draft_pick`. Each step
adds a word and narrows the grain. That's the grammar's "general to
specific, left to right" rule made visible. When you find yourself
wanting something *between* two of these, the name you want is usually
*also* between them alphabetically and conceptually.

## Recipe 8: A whole season at once, the fast way

**The story.** You don't want one game. You want *every* game --
a full season of play-by-play to model on.

Calling `espn_nba_pbp()` in a loop over a thousand `game_id`s would
work, and it would also be slow and rude to ESPN's servers. The `load_`
prefix exists precisely for this: pre-built, pre-cleaned season files.

```{r}
# load / nba / pbp. One call, one season, ~600k rows.
pbp_2024 <- load_nba_pbp(seasons = 2024)

# The whole family loads the same way.
load_nba_player_box(seasons = 2024)
load_nba_team_box(seasons = 2024)
load_nba_schedule(seasons = 2024)
```

The grammar lesson: the `load_` prefix changes the *performance
contract*, not the subject. `load_nba_pbp()` and `espn_nba_pbp()` both
hand you play-by-play; the first hands you a whole season fast, the
second hands you one game live. Reach for `load_` when you're doing
research across many games, and for `espn_` when you want one specific,
fresh game.

## Recipe 9: Pour it straight into a database

**The story.** A season of play-by-play is big. You'd rather not hold
it all in memory -- you want it in a database you can query.

Every `load_` function takes the same optional plumbing arguments, and
once you've seen them once you've seen them everywhere:

```{r}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "nba.sqlite")

# `dbConnection` + `tablename` -> write straight to disk while loading.
load_nba_pbp(
  seasons      = 2020:2024,
  dbConnection = con,
  tablename    = "nba_pbp"
)

dbDisconnect(con)
```

Margin note: consistency *is* a feature. Because every `load_` function
shares the `(seasons, ..., dbConnection, tablename)` signature, learning
one teaches you all of them. When a package keeps signatures parallel
like this, lean on it -- guess the arguments the same way you guess the
names.

## Working through a proxy

If you're on a corporate or campus network, your traffic may need to go
through a proxy. `hoopR` understands proxies in three layers, and you
should reach for the *least* invasive one that works.

**Layer 1 -- set it once for the whole session.** This is what you want
99% of the time. It covers every function in the package -- ESPN,
NBA Stats, KenPom, loaders, all of it.

```{r}
# Plain proxy.
options(hoopR.proxy = "http://proxy.company.com:8080")

# Authenticated proxy -- pass a list and it's spread into the request.
options(hoopR.proxy = list(
  url      = "http://proxy.company.com",
  port     = 8080,
  username = "saiem",
  password = Sys.getenv("PROXY_PASS")
))

# From here on, nothing changes -- just call functions normally.
espn_nba_team_roster(team_id = 13, season = most_recent_nba_season())
nba_leagueleaders(season = "2024-25")
```

**Layer 2 -- one call, one proxy.** Every NBA Stats API function
(`nba_*`) threads `...` down to the HTTP layer, so you can override the
proxy for a single call:

```{r}
nba_leaguestandings(
  season = "2024-25",
  proxy  = "http://other-proxy.company.com:3128"
)
```

A grammar-flavored caveat: per-call `proxy =` works for the `nba_*`
Stats API functions because they pass `...` through. The `espn_*`,
`kp_*`, and `load_*` functions don't take a per-call proxy -- for those,
use the session option in Layer 1. The prefix tells you the capability:
`nba_` functions are the ones wired for per-call overrides.

**Layer 3 -- let the environment decide.** If you'd rather not put a
proxy in your R code at all, `hoopR` will honor the standard
environment variables:

```{r}
Sys.setenv(
  http_proxy  = "http://proxy.company.com:8080",
  https_proxy = "http://proxy.company.com:8080",
  no_proxy    = "localhost,127.0.0.1"
)
```

These are read by libcurl automatically whenever you *haven't* set an
explicit proxy. It's the right layer for shared scripts and CI, where
the proxy is a property of the machine, not the analysis.

## Where to go next

You now have the grammar. Every recipe above was really the same trick:
decide where the data lives (the prefix), name the league, then name
the thing from general to specific. The NBA surface in `hoopR` is large,
but it is *regular* -- and regular things are guessable.

When you want the men's college game, the same grammar carries over
almost unchanged -- swap `nba` for `mbb`, meet a new prefix (`kp_` for
KenPom), and keep cooking. That's the MBB cookbook.
