---
title: "NBA Stats API Parameter Descriptions"
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://raw.githubusercontent.com/sportsdataverse/hoopR-mbb-data/main/themes/hoopR_gh.png"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
description: "A searchable reference table of every query parameter used by the NBA Stats API endpoints in <code>hoopR</code>, with accepted values and defaults."
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>"
)
with_dt <- requireNamespace("DT")
```

```{r eval = with_dt}
DT::datatable(hoopR::parameter_descriptions,options = list(scrollX = TRUE, pageLength = 25), filter = "top", rownames = FALSE)
```

```{r eval = !with_dt}
knitr::kable(hoopR::parameter_descriptions)
```

# **Our Authors**
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>

## **Our Contributors**
-   [Jason Lee](https://x.com/theFirmAISports)
<a href='https://x.com/theFirmAISports' target='blank'><img src='https://img.shields.io/twitter/follow/theFirmAISports?color=blue&amp;label=%40theFirmAISports&amp;logo=x&amp;style=for-the-badge' alt='@theFirmAISports'/></a>
<a href='https://github.com/papagorgio23' target='blank'><img src='https://img.shields.io/github/followers/papagorgio23?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@papagorgio23'/></a>
-   [Billy Fryer](https://x.com/BillyFryer42)
<a href='https://x.com/BillyFryer42' target='blank'><img src='https://img.shields.io/twitter/follow/BillyFryer42?color=blue&amp;label=%40BillyFryer42&amp;logo=x&amp;style=for-the-badge' alt='@BillyFryer42'/></a>
<a href='https://github.com/billyfryer' target='blank'><img src='https://img.shields.io/github/followers/billyfryer?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@billyfryer'/></a>
-   [Ross Drucker](https://x.com/rossdrucker9)
<a href='https://x.com/rossdrucker9' target='blank'><img src='https://img.shields.io/twitter/follow/rossdrucker9?color=blue&amp;label=%40rossdrucker9&amp;logo=x&amp;style=for-the-badge' alt='@rossdrucker9'/></a>
<a href='https://github.com/rossdrucker' target='blank'><img src='https://img.shields.io/github/followers/rossdrucker?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@rossdrucker'/></a>
-   [Vladislav Shufinskiy](https://x.com/vshufinskiy)
<a href='https://x.com/vshufinskiy' target='blank'><img src='https://img.shields.io/twitter/follow/vshufinskiy?color=blue&amp;label=%40vshufinskiy&amp;logo=x&amp;style=for-the-badge' alt='@vshufinskiy'/></a>
<a href='https://github.com/shufinskiy' target='blank'><img src='https://img.shields.io/github/followers/shufinskiy?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@shufinskiy'/></a>

## **Citation**

To cite the [**`hoopR`**](https://hoopR.sportsdataverse.org/) R package in publications, use:

BibTeX Citation

```bibtex
@misc{hoopr,
  author = {Saiem Gilani},
  title = {hoopR: Access Men’s Basketball Play by Play Data.},
  url = {https://hoopR.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
